---
title: "R Practice Problems - factors, lists, matrices"
author: "Ben Soibam"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Factors

##### 1. If `x <- factor(c("high", "low", "medium", "high", "high", "low", "medium"))`, write an R expression that will provide unique levels of x:
```{r,collapse=TRUE}
x <- factor(c("high", "low", "medium", "high", "high", "low", "medium"))
levels(x)
```

##### 2. Replacing a level of a factor will change contents of the factor that matches the level 
```{r,collapse=TRUE}
z <- factor(c("p", "q", "p", "r", "q"))
levels(z)
levels(z)[1] <- "w"
z
```

##### 3. If `s1 <- factor(sample(letters, size=5, replace=TRUE))`, and `s2 <- factor(sample(letters, size=5, replace=TRUE))`, write an R expression that will concatenate s1 and s2 in a single factor with 10 elements.
```{r,collapse=TRUE}
s1 <- factor(sample(letters, size=5, replace=TRUE))
s2 <- factor(sample(letters, size=5, replace=TRUE))
s <- factor(c(as.character(s1),as.character(s2)))
s
#Don't simply do the following. it will simply concatenate the numerical values of the levels.
ss <- c(s1,s2)
ss
```
## Lists

##### 4. Create a list called x with two elements; two vectors of length 1 called a and b whose value is 1 and 2 respectively.Add a third vector of length 1 to x, z = 3.
```{r,collapse=TRUE}
x <- list(a=1,b=2)
x
x[["z"]] <- c(3) 
x
```

##### 5. A list containing a vector, a matrix and a list
```{r,collapse=TRUE}
list_data <- list(c("Red","Green","Black"), matrix(c(1,3,5,7,9,11), nrow = 2),
list("Python", "PHP", "Java"))
list_data
```

##### 6.Converting a list to a vector using `unlist`

```{r,collapse=TRUE}
lst1 <- list(a=c(1,2,3),b=c(2))
lst2 <- list(a=c(1,2,3),b=c(TRUE,FALSE),c=c("this","is"))
un_lst1 <- unlist(lst1)
un_lst2 <- unlist(lst2)
un_lst1
typeof(un_lst1)
un_lst2
typeof(un_lst2)
lst3 <- list(a=list(1,2),b=c(TRUE,FALSE),c=c("this","is"))
un_lst3 <- unlist(lst3)
un_lst3
typeof(un_lst3)
```

## Matrices

##### 7. Write an R code to convert to a list where each column of matrix becomes a vector.

```{r,collapse=TRUE}
x = matrix(1:12, ncol=3)
x
print("list from the said matrix:")
l =  list(col1=x[,1],col2=x[,2],col3 = x[,3])
l
typeof(l)
```

##### 8. Write an R code to convert to a  matrix to a vector

```{r,collapse=TRUE}
x = matrix(1:12, ncol=3)
x
l = c(x)
l
typeof(l)
```


##### 9. Average for each row and each column in a matrix using `rowSums` and `colSums`

```{r,collapse=TRUE}
x = matrix(1:12, ncol=3)
x
print("row sums")
rowSums(x)
print("column sums")
colSums(x)
```

##### 10. using `order` function in R, sort a matrix based on one column

If `x <- c(10,1,30)` is a vector, `order(x)` gives the order of indices in `x` so that the items in x will be ascending (or descending) order.

For example, `order(x)` will give `2,1,3`; meaning items with indices 2, 1, and 3 are in increasing order. That means `x[order(x)]` will return a vector `(1,10,30)`.
```{r,collapse=TRUE}
x = matrix(sample(1:50,16), nrow=4, ncol=4)
x
```
Sort all the columns in matrix based on ascending order of the items in the first column

```{r,collapse=TRUE}
x[order(x[,1]),]
```





